Sustainable computational science:\\ the ReScience initiative

### 2017-11-09 replies_to_reviewers.md

# Replies to the reviewers' comments

## In reply to reviewer 1

We thank the reviewer for the in-depth examination of our work and for his useful suggestions for the future evolution of ReScience.

Concerning what the reviewer calls a defeatist attitude towards computational reproducibility, that is not at all how we see our approach. The manuscript focuses on reporting our experiences with ReScience. Concerning reproducibility, we note that having it verified by reviewers is not sufficient. We are not aware of any practical experience demonstrating that some other approach is efficient, which is why we consider the reproducibility of software environments an unsolved problem. Many ideas are being explored at the moment, including containerization as the reviewer mentions. The effectiveness of these ideas cannot be judged yet. Nobody knows, for example, if five years from now, the then-current version of Docker will still be able to handle today's container images. We see our attitude not as defeatist, but as prudent. Many of the authors are contributing to the development and evaluation of new ideas, and we will certainly try some of them out in the context of ReScience.

We have reworded some of our statements to make clearer that our doubts about reproducibility apply only to today's state of the art. We have also added a new paragraph on the reproducibility of software environments to the Outlook section.

### 2017-11-09 draft.tex
Some more details on conda

A fundamental technical issue is the difficulty of archiving or
accurately describing the software environments in which computational
scientists perform their work. A publication should be accompanied by
both a human-readable description of this environment and an
executable binary form. The human-readable description allows an
inspection of the versions of all software packages that were used,
for example to check for the impact of bugs that become known only
after a study was published. The executable version enables other
scientists to re-run the analyses and inspect intermediate
results. Ideally, the human-readable description would permit
rebuilding the executable version, in the same way that software
source code permits rebuilding executable binaries. This approach is
pursued for example by the package manager Guix \citep{Courtes:2015}.
A more limited but still useful implementation of the same idea exists
in the form of the conda package manager
\citep{conda}. Containerization, as implemented e.g. by Docker
\citep{docker} is currently much discussed, but provides only the
executable version without a human-readable description. Moreover, the
long-term stability of the container file format remains to be
evaluated. History has shown that long-term stability in computing
technology is achieved only by conventions for which it is a design
priority, for example the Java Virtual Machine \citep{JVM}. Docker, on
the contrary, is developed as a deployment technology with no ambition
towards long-term archiving.

>>>>

A fundamental technical issue is the difficulty of archiving or
accurately describing the software environments in which computational
scientists perform their work. A publication should be accompanied by
both a human-readable description of this environment and an
executable binary form. The human-readable description allows an
inspection of the versions of all software packages that were used,
for example to check for the impact of bugs that become known only
after a study was published. The executable version enables other
scientists to re-run the analyses and inspect intermediate
results. Ideally, the human-readable description would permit
rebuilding the executable version, in the same way that software
source code permits rebuilding executable binaries. This approach is
pursued for example by the package manager Guix \citep{Courtes:2015}.
A more limited but still useful implementation of the same idea exists
in the form of the conda package manager \citep{conda}, which uses a
so-called environment file to describe and reconstruct environments.
The main limitation compared to Guix is that the packages that make up
a conda environment are themselves not reproducible. For example, a
conda environment file does not state which compiler versions were
used to build a package.

Containerization, as implemented e.g. by Docker \citep{docker} is
currently much discussed, but provides only the executable version
without a human-readable description. Moreover, the long-term
stability of the container file format remains to be
evaluated. History has shown that long-term stability in computing
technology is achieved only by conventions for which it is a design
priority, for example the Java Virtual Machine \citep{JVM}. Docker, on
the contrary, is developed as a deployment technology with no ambition
towards long-term archiving.
